Sensitivity Analysis

### 2024-06-10 Article_Sensitivity_Analsysis_short-Chamomile.tex
10.06

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
This study aimed to investigate the supercritical extraction process of caraway oil from chamomile flowers. The distributed-parameter model describes the fluid-solid extraction process. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent are estimated from the Peng-Robinson equation of state. A set of laboratory experiments was performed under multiple constant operating conditions: $30 - 40^\circ C$, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. Sensitivity analyses play a crucial role in assessing the robustness of the findings or conclusions based on mathematical model. The local sensitivity analysis investigates the influence of infinitely small changes in the inlet temperature, pressure, and flow rate on the extraction yield.

>>>>

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
This study aims to investigate the extraction process of caraway oil from chamomile flowers. A mathematical model is used to describe the governing mass transfer phenomena in a solid-fluid environment under supercritical conditions using carbon dioxide as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent are estimated from the Peng-Robinson equation of state. A set of laboratory experiments was performed under multiple constant operating conditions: $30 - 40^\circ C$, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. Different sensitivity analysis methods can be applied to assess the robustness of the model parameters and its influence on the process model. The local sensitivity analysis investigates the influence of infinitely small changes in the model parameters and the controls on model output. In particular,  this work focuses on analysing the step change of pressure on the extraction yield and the model state space.

### 2024-06-10 Sections/Literature_Review.tex
10.06

	This study aims to analyse the influence of changes in operating conditions on the supercritical extraction process described in {\color{red}article 1}. The emphasis is put on the effect of the mass flow rate, pressure, and the inlet temperature. The relation between input and output is obtained by applying a sensitivity analysis. Sensitivity analysis examines the impact of varying inputs or model parameters on the system's output. The aim is to understand and allocate the source of uncertainty in the output to the corresponding inputs or parameters. There are many sensitivity analysis methods, which include but are not limited to those listed below:

@@

	Different supercritical extraction models were analysed using sensitivity analysis. \citet{Fiori_2007}, performed the sensitivity calculations by varying the parameters within their confidence interval and observing how the model results changed. This allows to evaluate the effect of the uncertainties on model predictions. The sensitivity analysis revealed that the particle diameter and the internal mass transfer coefficient are significant for the extraction process. The effect of changing some operative conditions was also investigated, underlining how the solvent flow rate and the seed milling affect the extraction process.

	\citet{Santos2000}, in their work, considered a model of supercritical extraction process for semi-continuous isothermal and isobaric extraction process using carbon dioxide as a solvent. The parametric sensitivity analysis was carried out by applying disturbances of 10\% in the values of the normal operation conditions.

	\citet{Hatami2024} used a one-factor-at-a-time sensitivity analysis to assess the response of NPV concerning variations in both technical and economic variables. Their findings show that the most influential factors on NPV include the price of the extract, the interest rate, the dynamic time of SFE, and the project lifetime.

	\citet{Poletto1996} provided a general dimensionless model for the supercritical extraction process of vegetable and essential oils and applied a sensitivity analysis. They found that a dimensionless partition coefficient and a dimensionless characteristic time appeared as the most important parameters of the extraction process. The sensitivity calculations were performed by varying the parameters and analysing the model response.

>>>>

	This study aims to analyse the influence of changes in operating conditions on the process model described in {\color{red}article 1}. The relation between input and output is obtained by applying a sensitivity analysis, which examines the impact of model parameters or controls on the model output. The result of sensitivity analysis can be used to allocate the source of uncertainty in the system, perform the model reduction or search for errors in the model (by identifying unexpected relationships between inputs and outputs). There are many sensitivity analysis methods, which include but are not limited to those listed below:

@@

	In the literature, many different supercritical extraction models were analysed using different sensitivity analysis techniques. \citet{Fiori_2007}, performed the sensitivity calculations by varying the parameters within their confidence interval and observing how the model results changed. Their sensitivity analysis revealed that changes in particle diameter and the internal mass transfer coefficient have the strongest influence on extraction in the second phase of the process (diffusion-control regime).% The effect of changing some operative conditions was also investigated, underlining how the solvent flow rate and the seed milling affect the extraction process.

	\citet{Santos2000}, in their work, considered the model of \citet{Sovova1994} for the semi-continuous isothermal and isobaric extraction processes using carbon dioxide as a solvent. The parametric sensitivity analysis was carried out using a factorial design in two levels. The model parameters were disturbed by 10\%, and their main effects were analysed and strategies for high-performance operation are proposed. The authors performed the sensitivity analysis with respect to the superficial velocity, particle diameter, initial concentration of solute in the solid phase and the concentration of solute in the fluid phase at the inlet to the extractor.

	\citet{Hatami2024} performed the one-factor-at-a-time sensitivity analysis to assess the response of net present value concerning variations in both technical and economic variables. Their study is divided into two parts. In the first part, the sensitivity analysis is carried out for an extraction plant with a constant extractor volume of 300 L. The focus is on examining how net present value is influenced by changes in individual technical and economic parameters while keeping the extractor volume fixed. In the second part, the investigation revolves around the effects of altering the extractor volume (from 1 to 600 L) on the project's overall profitability. Their findings show that the most influential factors on net present value include the price of the extract, the interest rate, the dynamic time, and the project lifetime.

	%\citet{Poletto1996} provided a general dimensionless model for the supercritical extraction process of vegetable and essential oils. The sensitivity calculations were performed by varying the parameters and analysing the model response. The authors found that a dimensionless partition coefficient and a dimensionless characteristic time appeared as the most important parameters of the extraction process.

### 2024-06-10 Sections/Model.tex
10.06

			If the value of enthalpy $h$ is known from the time evolution of the energy equation, and pressure $P$ is known from measurement, then the temperature $T$ can be reconstructed based on the departure function. The departure function is a mathematical function that characterizes the deviation of a thermodynamic property (enthalpy, entropy, and internal energy) of a real substance from that of an ideal gas at the same temperature and pressure. The departure function is defined as the difference between the value of a thermodynamic property for a real fluid and the corresponding value for an ideal gas at the same temperature and pressure. They are computed by integrating a function that depends on the equation of state and its derivatives. As presented by \citet{Gmehling2019}, for the Peng-Robinson equation of state, the enthalpy departure function is defined by Equation \ref{EQ:Enthalpy_PR}.

@@

			Equation \ref{EQ:Enthalpy_PR} requires an reference sate, which is assumed to be $T_{ref}=298.15$[K] and $P_{ref}=1.01325$[bar].

			A root-finder can be used to find a value of temperature, which minimizes the difference between the value of enthalpy coming from the heat balance and the departure functions. The root fining procedure to repeated at every time step to find a local temperature along spatial direction $z$.

>>>>

			If the value of enthalpy $h$ is known from the time evolution of the energy equation, and pressure $P$ is known from measurement, then the temperature $T$ can be reconstructed based on the departure function. The departure function is a mathematical function that characterizes the deviation of a thermodynamic property (enthalpy, entropy, and internal energy) of a real substance from that of an ideal gas at the same temperature and pressure. As presented by \citet{Gmehling2019}, for the Peng-Robinson equation of state, the enthalpy departure function is defined by Equation \ref{EQ:Enthalpy_PR}.

@@

			Equation \ref{EQ:Enthalpy_PR} requires an reference sate, which is assumed to be $T_{ref}=298.15$ K and $P_{ref}=1.01325$ bar.

			A root-finder can be used to find a value of temperature, which minimizes the difference between the value of enthalpy coming from the heat balance and the departure functions. The root fining procedure to repeated at every time step to find a temperature profile along spatial direction $z$.

### 2024-06-10 Sections/Results_Sensitivity.tex
10.06

	This study examines the impact of pressure change on the state space and extraction yield. The process model and parameters are described in {\color{red}article 1}. Calibration of the process model is based on experiments performed under varying operating conditions: temperatures between $30 - 40^\circ C$, pressures between $100 - 200$ bar, and mass flow between $3.33-6.67 \cdot 10^{-5}$ kg/s. The local sensitivity analysis was conducted, assuming the system operates at the midpoint of the validated range: $35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s.

>>>>

	This study examines the impact of pressure change on the state space and extraction yield. The process model and parameters are described in {\color{red}article 1}. Calibration of the process model is based on experiments performed under multiple  operating conditions: temperatures between $30 - 40^\circ C$, pressures between $100 - 200$ bar, and mass flow rates between $3.33-6.67 \cdot 10^{-5}$ kg/s. The local sensitivity analysis was conducted, assuming the system operates at the midpoint of the validated range: $35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s.

### 2024-06-10 Sections/Sensitivity_Analysis.tex
10.06

	The sensitivity matrix $S(x;\Theta)$ represents the matrix of equations of size $N_x \times N_\Theta$, where each subequation $S(n_x,n_\Theta)$ is the derivative of the state variable $x_{n_x}$ with respect to the parameter $\Theta_{n_\Theta}$. Matrix ${J_x}(x;\Theta)$ and $S(x;\Theta)$ describe indirect influence of $\Theta_{n_\Theta}$ on the model output.

@@

	The Jacobian ${J_\Theta}(x;\Theta)$ represents the matrix of equations of size $N_x \times N_\Theta$, where each subequation ${J_\Theta}(n_x,n_\Theta)$ is the partial derivative of the process model equation $G_{n_x}$ with respect to the parameter $\Theta_{n_\Theta}$. ${J_\Theta}(n_x,n_\Theta)$ defines direct effect of $\Theta_{n_\Theta}$ on the model output.

@@

	The combined system containing the original set of equations ${G}(x;\Theta)$ and sensitivity equations can be formulated as ${\textbf{G}}\left(x;\Theta\right)$. The size of ${\textbf{G}}\left(x;\Theta\right)$ is equal to $N_s = N_x(N_\Theta + 1)$.

>>>>

	The sensitivity matrix $S(x;\Theta)$ represents the matrix of equations of size $N_x \times N_\Theta$, where each subequation $S(n_x,n_\Theta)$ is the derivative of the state variable $x_{n_x}$ with respect to the parameter $\Theta_{n_\Theta}$. Matrix ${J_x}(x;\Theta)$ and $S(x;\Theta)$ describe indirect influence of $\Theta_{n_\Theta}$ on the state space.

@@

	The Jacobian ${J_\Theta}(x;\Theta)$ represents the matrix of equations of size $N_x \times N_\Theta$, where each subequation ${J_\Theta}(n_x,n_\Theta)$ is the partial derivative of the process model equation $G_{n_x}$ with respect to the parameter $\Theta_{n_\Theta}$. ${J_\Theta}(n_x,n_\Theta)$ defines direct effect of $\Theta_{n_\Theta}$ on the state space.

@@

	The augmented system containing the original set of equations ${G}(x;\Theta)$ and sensitivity equations can be formulated as ${\textbf{G}}\left(x;\Theta\right)$. The size of ${\textbf{G}}\left(x;\Theta\right)$ is equal to $N_s = N_x(N_\Theta + 1)$.
